---
title: "P8105 Homework2"
author: Yushan Wang
output: github_document
---

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(ggridges)
library(readxl)
library(dplyr)
```

## Probelm 1

Clean the Mr.trash wheel dataset by omit non-data & note entries, omit non-dumpster-specific rows, give reasonable names, and round sports balls to integer

```{r}
# import mr.trash dataset and omit note entries
mr_trashwheel_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", 
             sheet = "Mr. Trash Wheel",
             range = "A2:N408") %>% 
  
  # omit non-data entries
  select(-Month) %>% 
  
  # omit non-dumpster-specific rows
  drop_na() %>% 
  
  # give reasonable names
  janitor::clean_names() %>% 
  
  # round sports balls
  mutate(sports_balls = round(sports_balls))
```

Import the 2018 & 2019 precipitation dataset

```{r}
# import 2018 precipitation dataset and omit note entries and NA entries
precipitation_2018 <-
  read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", 
             sheet = "2018 Precipitation",
             range = "A2:B14")  

# import 2019 precipitation dataset and omit note entries and NA entries
precipitation_2019 <-
  read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", 
             sheet = "2019 Precipitation",
             range = "A2:B14")  
```

Combine two dataframe, add a new variable "year", and convert month number to character

```{r}
# create function to convert month number to character
my.month.name <- Vectorize(function(n) c("january", "february", "narch", 
                                         "april", "may", "june", "july", 
                                         "august", "september", "october",
                                         "november", "december")[n])

precipitation_df <- 
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  mutate("year" = c(rep(2018, 12), rep(2019,12) ),
         Month = my.month.name(Month)) %>% 
  relocate(year) %>% 
  janitor::clean_names()

```

**Descriptions of the above datasets**

There are 344 observations in Mr.trashwheel dataset.
Variable "dumpster" is denoted by numbers range from 1 to 344.
Variable "date" is coded as year-month-day. For example, `r mr_trashwheel_df[3,3]`.
All other variables are coded as numbers.
 
There are 18 observations in combined 2018 & 2019 precipitation dataset.
Variable "year" consists of 2018 and 2019.
Variable "month" consists of all twelve month in character
The total precipitation, denoted by variable "total", range from `r min(pull(precipitation_df, total))` to `r max(pull(precipitation_df, total))`

Total precipitation in 2018 was `r sum(subset(precipitation_df, year == 2018)$total)`.

The median number of sports balls in a dumpster in 2017 was `r median(subset(mr_trashwheel_df, year == 2017)$sports_balls)`.

## Probelm 2

Clean the data in pols-month.csv
  
```{r}
pols_month_df <- 
  
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), "-") 

#remove day, prez_dem and prez_gop
pols_month_df <- 
  subset(pols_month_df, select = -c(day, prez_dem, prez_gop)) %>% 
  
  # convert month to name use function defined in problem 1
  mutate(
    month = str_remove(month, "^0+"), # remove leading 0
    month = as.numeric(month), 
    month = my.month.name(month)
    ) %>% 
  
  pivot_longer(
    gov_gop:rep_dem,
    names_to = "president",
    #names_prefix = "gov_"
    values_to = "pol_value"
  ) 

```

Clean the data in snp.csv

```{r}
snp_df <- 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), "/") %>% 
  rename(snp = close) %>% 
  mutate(
    month = as.numeric(month),
    year = as.numeric(year),
    year = ifelse(year > 90, year + 1900, year + 2000),
    month = my.month.name(month),
    ) %>% 
  relocate(year, month, day)
  

#remove day 
snp_df <- subset(snp_df, select = -day)

```

Clean the unemployment data

```{r}

unemployment_df <- 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  na.omit(unemployment_df) #omit NA rows

# renaming month
unemployment_df <- rename(unemployment_df, 
                          january = jan,
                          february = feb,
                          march = mar,
                          april = apr,
                          may = may,
                          june = jun,
                          july = jul,
                          august = aug,
                          september = sep,
                          october = oct,
                          november = nov,
                          december = dec) %>% 
  
  pivot_longer(
    january:december,
    names_to = "month",
    values_to = "unemployment"
  ) 
  
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
pol_snp <- merge(pols_month_df, snp_df , by = c("year", "month")) 
pol_snp_unemployment <- merge(pol_snp, unemployment_df , by = c("year", "month")) 

```

**Description of dataset**

The pols_month_df dataset has `r dim(pols_month_df)[1]` rows and  `r dim(pols_month_df)[2]` columns. The years in this dataset range from `r min(pull(pols_month_df, year))` to `r max(pull(pols_month_df, year))`. The variable "president" include character gop and dem, and the variable pol_value range from `r min(pull(pols_month_df, pol_value))` to `r max(pull(pols_month_df, pol_value))`.

The snp_df dataset has `r dim(snp_df)[1]` rows and  `r dim(snp_df)[2]` columns. The years in this dataset range from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. The variable "snp" range from `r min(pull(snp_df, snp))` to `r max(pull(snp_df, snp))`.

The unemployment_df dataset has `r dim(unemployment_df)[1]` rows and  `r dim(unemployment_df)[2]` columns. The years in this dataset range from `r min(pull(unemployment_df, year))` to `r max(pull(unemployment_df, year))`. The variable "snp" range from `r min(pull(unemployment_df, unemployment))` to `r max(pull(unemployment_df, unemployment))`.

## Probelm 3

Clean the Popular_Baby_Names.csv

```{r}
babynames_df <- 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  
  # remove repeated rows
  distinct(ethnicity, childs_first_name, .keep_all = TRUE) %>% 
  
  # convert every word in dataframe lowercase
  mutate(gender = tolower(gender),
         ethnicity = tolower(ethnicity),
         childs_first_name = tolower(childs_first_name)
          )

# show the resulting datafram
babynames_df
```
Produce table showing the rank in popularity of the name “Olivia” as a female baby name over time

```{r}
olivia_df = filter(babynames_df, childs_first_name == "olivia", gender == "female") 
olivia_df = subset(olivia_df, select = -c(gender, childs_first_name, count))

# show the resulting datafram
olivia_df
```
Produce table showing the most popular name among male children over time.

```{r}
popular_df = filter(babynames_df, rank == "1", gender == "male") 
popular_df = subset(popular_df, select = -c(gender, rank, count))

# show the resulting datafram
popular_df
```

Scatter Plot

```{r}
male_white_nonhis = filter(babynames_df, year_of_birth == 2016, 
                         ethnicity == "white non hispanic", 
                         gender == "male") 

male_white_nonhis %>% 
  ggplot(aes(x = rank, y = childs_first_name)) + 
  geom_point(alpha = .5)



```






